---
title: "project_data"
author: "Meng Luo"
date: "2021/11/14"
output: html_document
---

```{r setup, include=F}

```
## Data
A series of data preprocessing steps were used to acquire the monthly and yearly area-averaged GPP data over ABZ. 

Firstly, used spatial masks of ABZ to clip and calculate the area-averaged GPP values for each data.

Secondly, unified the GPP units of climate models and remote sensing products.

Thirdly, filtered the GPP data from different data source to the overlapping time periods from 2003 to 2012.

Fourthly, aggregated the monthly GPP to annual scales.

Fifthly, the ensemble mean GPP values from four remote sensing products was calculated. Note that GISS in CMIP5 and UKESM in CMIP6 only covers 2006-2012. 

## Code

```{r read NC, include = T}
#install.packages("raster")
library(cowplot)
library(raster)
library(ncdf4)
library(rasterVis)
library(lattice)
library(tidyverse)
library(rgdal)
library(Directional)
library(dplyr)
library(reshape2)




##read Remote sensing GPP product#######################################################
#lct <- Sys.getlocale("LC_TIME"); Sys.setlocale("LC_TIME", "C")
## GPP unit: g m-2 d-1##################################################################
RS_GPP = read.table("D:/Research/GCAM/Code/STAT_433/remotely_sensed_GPP.csv",header = TRUE, sep = ",")

#RS_GPP1: MODIS,RS_GPP2:BESS, RS_GPP3:VPM, RS_GPP4:FLUXCOM

RS_GPP_1 = RS_GPP %>% 
  group_by(year,month)%>%
  summarise(MODIS_mean = mean(RS_GPP1, na.rm=TRUE),BESS_mean = mean(RS_GPP2, na.rm=TRUE) , VPM_mean = mean(RS_GPP3, na.rm=TRUE) ,FLUXCOM_mean = mean(RS_GPP4, na.rm=TRUE)) %>%
  mutate(year_month = year*100+month)
# remove nan
RS_GPP_1 = na.omit(RS_GPP_1)





##read model GPP########################################################################
## GPP unit kg m-2 s-1###############################################
# CMIP5 ###############################################################################
#######################################################################################
#GPP data1#######################################################
ncfile5_1<-nc_open("D:/PHD/Course/STAT 433/project/data/CMIP5/new/gpp_Lmon_GISS-E2-R_historicalExt_r6i1p1_200601-201212.nc")
GISS5 = ncvar_get(ncfile5_1, varid='gpp', start=c(1,1,1), count=c(-1,-1,84), verbose=FALSE, signedbyte=TRUE, collapse_degen=TRUE)
# GPP data2######################################################
ncfile5_2<-nc_open("D:/PHD/Course/STAT 433/project/data/CMIP5/new/same model/gpp_Lmon_IPSL-CM5A-MR_historicalNat_r3i1p1_185001-201212.nc")
IPSL5 = ncvar_get(ncfile5_2, varid='gpp', start=c(1,1,1801), count=c(-1,-1,156), verbose=FALSE, signedbyte=TRUE, collapse_degen=TRUE)
# GPP data3######################################################
ncfile5_3<-nc_open("D:/PHD/Course/STAT 433/project/data/CMIP5/new/same model/gpp_Lmon_CanESM2_historicalMisc_r1i1p2_185001-201212.nc")
CanESM5 = ncvar_get(ncfile5_3, varid='gpp', start=c(1,1,1801), count=c(-1,-1,156), verbose=FALSE, signedbyte=TRUE, collapse_degen=TRUE)






# CMIP6 ###############################################################################
#######################################################################################
#GPP data1###########################################################
ncfile6_1<-nc_open("D:/PHD/Course/STAT 433/project/data/CMIP6/GPP/select/gpp_Lmon_UKESM1-0-LL_hist-piAer_r1i1p1f2_gn_195001-201412.nc")
UKESM16 = ncvar_get(ncfile6_1, varid='gpp', start=c(1,1,601), count=c(-1,-1,180), verbose=FALSE, signedbyte=TRUE, collapse_degen=TRUE)
# GPP data2###########################################################
ncfile6_2<-nc_open("D:/PHD/Course/STAT 433/project/data/CMIP6/GPP/select/gpp_Lmon_IPSL-CM6A-LR_historical_r10i1p1f1_gr_185001-201412.nc")
IPSL6 = ncvar_get(ncfile6_2, varid='gpp', start=c(1,1,1801), count=c(-1,-1,180), verbose=FALSE, signedbyte=TRUE, collapse_degen=TRUE)
# GPP data3###########################################################
ncfile6_3<-nc_open("D:/PHD/Course/STAT 433/project/data/CMIP6/GPP/select/gpp_Lmon_CanESM5_historical_r1i1p2f1_gn_185001-201412.nc")
CanESM6 = ncvar_get(ncfile6_3, varid='gpp', start=c(1,1,1801), count=c(-1,-1,180), verbose=FALSE, signedbyte=TRUE, collapse_degen=TRUE)





#ABOVE raster################################################################################################
#ABoVE raster1#################################################
ABoVE_raster1 <- raster(x = "D:/PHD/Course/STAT 433/project/data/ABoVe_region/above_domain2_192_144.tif")
#plot(ABoVE_raster1)
ABoVE_raster1 = (as.matrix(ABoVE_raster1))
ABoVE_raster1 = t(apply(ABoVE_raster1, 2, rev))
ABoVE_raster1[ABoVE_raster1 == 0] = NaN
#ABoVE raster2#################################################
ABoVE_raster2 <- raster(x = "D:/PHD/Course/STAT 433/project/data/ABoVe_region/above_domain2_143_144_360.tif")
#plot(ABoVE_raster2)
ABoVE_raster2 = (as.matrix(ABoVE_raster2))
ABoVE_raster2 = t(apply(ABoVE_raster2, 2, rev))
ABoVE_raster2[ABoVE_raster2 == 0] = NaN
#ABoVE raster3#################################################
ABoVE_raster3 <- raster(x = "D:/PHD/Course/STAT 433/project/data/ABoVe_region/above_domain2_64_128_360.tif")
#plot(ABoVE_raster3)
ABoVE_raster3 = (as.matrix(ABoVE_raster3))
ABoVE_raster3 = t(apply(ABoVE_raster3, 2, rev))
ABoVE_raster3[ABoVE_raster3 == 0] = NaN
#ABoVE raster4#################################################
ABoVE_raster4 <- raster(x = "D:/PHD/Course/STAT 433/project/data/ABoVe_region/above_domain2_90_144.tif")
#plot(ABoVE_raster4)
ABoVE_raster4 = (as.matrix(ABoVE_raster4))
ABoVE_raster4 = t(apply(ABoVE_raster4, 2, rev))
ABoVE_raster4[ABoVE_raster4 == 0] = NaN





# calculate mean GPP####################################################################
GISS_mean5 = rep(NaN, 180)
IPSL_mean5 = rep(NaN, 180)
CanESM_mean5 = rep(NaN, 180)

UKESM1_mean6 = rep(NaN, 180)
IPSL_mean6 = rep(NaN, 180)
CanESM_mean6 = rep(NaN, 180)

year = rep(NaN, 180)
month = rep(NaN, 180)

tran = 3600*24*1000

for(time_i in 1:180)
{
  if (time_i%%12 == 0)
  {year[time_i] = time_i%/%12 + 1999;

   month[time_i] =  12;}
  else
   {year[time_i] = time_i%/%12 + 2000;

   month[time_i] =  time_i%%12;
   }
# unit transform at the same time
# unit: g m-2 d-1
  
  IPSL_mean6[time_i] = mean(ABoVE_raster2*IPSL6[,,time_i]*tran, na.rm = TRUE)
  CanESM_mean6[time_i] = mean(ABoVE_raster3*CanESM6[,,time_i]*tran, na.rm = TRUE)
  if(time_i<157)
  {
    IPSL_mean5[time_i] = mean(ABoVE_raster2*IPSL5[,,time_i]*tran, na.rm = TRUE)
    CanESM_mean5[time_i] = mean(ABoVE_raster3*CanESM5[,,time_i]*tran, na.rm = TRUE)
  }
  if(time_i>72 & time_i<157)
  {
    GISS_mean5[time_i] = mean(ABoVE_raster4*GISS5[,,time_i-72]*tran, na.rm = TRUE)
    UKESM1_mean6[time_i] = mean(ABoVE_raster1*UKESM16[,,time_i]*tran, na.rm = TRUE)
    }
}


GPP_data<-data.frame(year,month,GISS_mean5,IPSL_mean5,CanESM_mean5,UKESM1_mean6,IPSL_mean6,CanESM_mean6)%>%
  mutate(year_month = year*100+month)

GPP_data_all= inner_join(GPP_data,RS_GPP_1)
#MODIS,RS_GPP2:BESS, RS_GPP3:VPM, RS_GPP4:FLUXCOM
GPP_data_all = GPP_data_all%>%
  mutate(GPP_RS_mean = (MODIS_mean+BESS_mean+VPM_mean+FLUXCOM_mean)/4)%>%
  filter(year<=2012)

saveRDS(GPP_data_all, file = "D:/PHD/Course/STAT 433/project/GPP_app/GPP_data_all3.rds")


```

Figure. Monthly time series of GPP from (A) climate models and (B) remote sensing observations.

